---
title: "Coding Project 5"
author: "Wanzhu Zheng"
date: "2023-06-07"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rvest)
library(xml2)
library(ggplot2)
library(dplyr)
library(magrittr)
library(raw)
library(stringr)
library(readr)
```

## Read Table from W3Schools

```{r}
html = read_html("https://www.w3schools.com/tags/ref_urlencode.ASP")
tbls_nodes = html_nodes(html, "table")
 
tbls_content = tbls_nodes %>% html_table(fill=TRUE)
tbls = as.data.frame(tbls_content[[1]])
tbl = tbls[c(1,3)]
head(tbl)
```

We notice that R cannot deal with particular %-encodings.\
For example:

```{r}
tbl[109:114, ]
```

We deal with this by omitting those rows.

```{r}
mapping =  tbl[-c(96, 110, 112, 113, 126, 129, 142), ]
head(mapping)
```

## Create Sample Strings

```{r}
source("preallocated.R")
source("vectorized.R")
source("my_funs.R")
source("words.R")
source("times.R")
```

We write a create_n_str function that samples letters, numbers, and
UTF-8. It takes in an integer n and outputs n sample strings.
```{r eval=FALSE}
create_n_str = function(n) {
  str_len = sample(1:200000, n, replace = TRUE)
  chars = sample(c(letters, 0:9, mapping$`From UTF-8`), sum(str_len), replace=TRUE)
  tapply(chars, rep(1:n, str_len), paste, collapse="")
}
```

As expected, when we input n = 1, the function returns 1 string and if
we input n = 25, the function returns 25 strings.

```{r}
length(create_n_str(1))
length(create_n_str(25))
```

## Run-time Performance as a Function of Size

We write another create_str function that samples letters, numbers, and
UTF-8.\
It takes in an integer n and outputs a string of length n.

```{r eval=FALSE}
create_str = function(n) {
  str_len = n
  chars = sample(c(letters, 0:9, mapping$`From UTF-8`), str_len, replace=TRUE)
  paste(chars, collapse="")
}
```

We want to see how URLdecode() compares to itself at different values of
n passed into create_str. We first create a list of sample strings

```{r}
words = readRDS("words.rds")
```

Then, we use system.time() to time the runtime of URLdecode() for
strings ranging from length 1 to 200K.

```{r}
tm0 = readRDS("tm0.rds")
```

Plot Size vs. Run-time

```{r}
tm0_df = data.frame(size=i, time=tm0)

ggplot(data=tm0_df, aes(x=size, y=time, group=1)) +
  geom_line(color="red4")+
  geom_point()
```

Notice that there is a quadratic relationship between string size and
runtime.

## Pre-allocated Version of URLdecode()

We modify the URLdecode() function to use preallocation which will give
us a faster runtime.

```{r eval=FALSE}
preallocated = function (URL) {
  vapply(URL, function(URL) {
    x = charToRaw(URL)
    pc = charToRaw("%")
    out = raw(length(x))
    i = 1L
    while (i <= length(x)) {
      if (x[i] != pc) {
        out[i] = x[i]
        i = i + 1L
      }
      else {
        y = as.integer(x[i+1L:2L])
        y[y > 96L] = y[y > 96L] - 32L
        y[y > 57L] = y[y > 57L] - 7L
        y = sum((y - 48L) * c(16L, 1L))
        out[i] = as.raw(as.character(y))
        i = i + 3L
      }
    }
    rawToChar(out[out != 00])
  }, character(1), USE.NAMES = FALSE)
}
```

### Verification

In this section, we verify that our output from preallocated() is the
same as our output from URLdecode(). In order to do so, we check that
preallocated() returns the same content as URLdecode() for all test
strings:

```{r}
for (i in words) {
  stopifnot(preallocated(i) == URLdecode(i))
}
```

Since no error was returned, we know that all three conditions have been
met and our preallocated() function returns the same character string as
URLdecode().

## Vectorized Version of URLdecode()

```{r eval=FALSE}
vectorized = function(URL) {
  x = charToRaw(URL)
  pc = charToRaw("%")
  out = raw(length(x))
  
  pcidx = unlist(gregexpr("%", URL))
  pct = rep(pcidx, each=2)
  
  pc_pair = pct + c(1,2)
  y = as.integer(x[pc_pair])
  y[y > 96L] = y[y > 96L] - 32L
  y[y > 57L] = y[y > 57L] - 7L
  
  proc = (y - 48L) * c(16L, 1L)
  m = matrix(proc, ncol=2, byrow=TRUE)
  sum = rowSums(m)
  out[pcidx] = as.raw(as.character(sum))
  
  non_pc = c(pcidx, pc_pair)
  nonidx = setdiff(1:length(x), non_pc)
  out[nonidx] = x[nonidx]
  
  return(rawToChar(out[out != 00]))
}
```

## Debugging
### Preallocated Version
* Warning: Exists 00 characters when converting to string
    + There was a warning message that told me my final output string had
    00 characters when I converted rawToChar() for my final output string.
    + In preallocation, we are passing %-encoded pairs into the index of
    the "%" in the string. Therefore, there will be lots of empty raw characters
    since the two indices after "%" will not be filled. I solved this by adding
    rawToChar(out[out != 00]) at the end of the function. This returns the
    indices in out that are not 00.

### Vectorized Version
* Warning: y = as.integer(x[pc_idx+1L:2L])
    + This gave a warning message stating that the list size was not correct.
    + I solved this by using the rep() function to replicate each
    value of pc_idx twice. Then, I added c(1, 2) to the indices. This gives me
    the next two indices of characters after %.
* Error: Incorrect sums
    + I wasn't able to figure out how to add the pairs together, and by calling
    the sum() function directly, it added all the values in the vector together.
    + I solved this by putting all my values in a matrix defined by 2 columns.
    The first column was the first index after the %, and the second column was
    the second index after the %. I called rowSums() on the matrix, and the 
    result was a list of summed pairs.
* Warning: setdiff(1:out, non_pc)
    + This gave a warning stating numerical expression has 52 elements, only 
    the first used.
    + I solved this by replacing out, which is a raw list, to length(x), which
    gives an integer. out has the same length as x, but was converted to raw
    in the beginning to store raw values.
    

## All Runtime Plots
We return all the times as a dataframe by size of input.
```{r}
tm1 = readRDS("tm1.rds")
tm2 = readRDS("tm2.rds")
tm1_df = data.frame(size=i, time=tm1)
tm2_df = data.frame(size=i, time=tm2)

tm_tbl = cbind(tm0_df$size, tm0_df$time, tm1_df$time, tm2_df$time)
colnames(tm_tbl) = c("size", "url time", "preallocated time", "vectorized time")
tm_df = as.data.frame(tm_tbl)

tm_df
```

Then, we plot:
```{r}
colors <- c("URLdecode()"="cadetblue4", 
            "preallocated()"="darkolivegreen", 
            "vectorized()"="pink3")

ggplot(tm_df, aes(x=`size`)) +
    geom_line(aes(y=`url time`, color="URLdecode() time")) +
    geom_line(aes(y=`preallocated time`, color="preallocated time")) +
    geom_line(aes(y=`vectorized time`, color="vectorized time")) +
    labs(x="size (n)",
         y="time (seconds)",
         color="legend") +
    scale_color_manual(values=colors)
```
As we can see, our vectorized and preallocated functions return linear time.
vectorized() outperforms URLdecode() and preallocated() since it has the
smallest slope. However, we can notice that preallocated() and vectorized()
are both very close in run-time.

## 591K %-Encoded String
```{r}
readRDS("url_591k.rds")
readRDS("pre_591k.rds")
readRDS("v_591k.rds")
```
My estimate of the run-time for URLdecode() was pretty accurate because
it shows a quadratic run-time, and the run-time for the 591K string falls on
the quadratic curve.  
The vectorized function is the fastest, as expected.

